---
title: "Coronavirus Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
`%>%` <- magrittr::`%>%`
library(flexdashboard)
source("./functions/dashboard_functions.R")
load("./data/gis_mapping.RData")
coronavirus_jhu = coronavirus::refresh_coronavirus_jhu()
gis_codes = get_gis_codes()
corona_agg <- coronavirus_agg(coronavirus_jhu = coronavirus_jhu,
                gis_codes = gis_codes)
df_cases2 <- coronavirus_jhu %>%
  dplyr::filter(location_type == "country",
                !is.na(location_code),
                data_type != "recovered_new") %>%
  dplyr::left_join(gis_codes %>% 
                     dplyr::select(-lat, - long) %>%
                     dplyr::filter(combined_key == country_region, 
                                               is.na(province_state)), by = c("location" = "combined_key")) %>%
  dplyr::left_join(continent_mapping %>% dplyr::select(continent_name, continent_code, iso2), by = "iso2")
  
head(df_cases2)
table(is.na(df_cases2$continent_name))
df_cases_c <- df_cases2 %>% 
  dplyr::filter(data_type == "cases_new",
                !is.na(continent_name)) %>%
  dplyr::group_by(date, continent_name) %>%
  dplyr::summarise(total = sum(value), .groups = "drop")

# Smoothing outliers
head(df_cases_c)
table(is.na(df_cases_c$total))
table(is.na(df_cases_c$continent_name))


df_cases_c$total[which(df_cases_c$date == as.Date("2021-05-20") & 
                         df_cases_c$continent_name == "Europe")] <- (df_cases_c$total[which(df_cases_c$date == as.Date("2021-05-19") & 
                                                                                              df_cases_c$continent_name == "Europe")] + 
                                                                       df_cases_c$total[which(df_cases_c$date == as.Date("2021-05-21") & 
                                                                                                df_cases_c$continent_name == "Europe")]) / 2

last_update <- max(df_cases_c$date)

```


<script>
document.querySelector(".navbar-header > span.navbar-brand > span.navbar-author").innerHTML = "Data Up-To-Date: `r last_update`";
</script>

<style>
.navbar-author {
  position: absolute;
  right: 1rem;
}
</style>

Summary
===================================== 

Row 
-----------------------------------------------------------------------

### Mortality Rate - Death / Confirmed Cases Ratio

```{r}
mortality_plot(coronavirus_agg = corona_agg)
```


### Confirmed Cases Distribution by Country

```{r}
d <- corona_agg %>% dplyr::mutate(parent = "Confirmed")

plotly::plot_ly(
  data = d,
  type= "treemap",
  values = ~confirmed,
  labels= ~ location,
  parents=  ~ parent,
  domain = list(column=0),
  #name = "Confirmed",
  textinfo="label+value+percent parent"
)  %>%
  plotly::layout(margin = list(r = 0, l = 0, t = 0, b = 0))

```


Row 
-----------------------------------------------------------------------


### Distribution of Confirimed Cases by Continent

```{r}
# Fix for South America negative values during Dec 25 2022
# Using average of previous and next day

df_cases_c$total[which(df_cases_c$date == as.Date("2022-12-25") & 
                   df_cases_c$continent_name == "South America")] <- (df_cases_c$total[which(df_cases_c$date == as.Date("2022-12-24") & 
                                                                                              df_cases_c$continent_name == "South America")] +
  df_cases_c$total[which(df_cases_c$date == as.Date("2022-12-26") & 
                           df_cases_c$continent_name == "South America")] ) / 2

plotly::plot_ly(data = df_cases_c,
                x = ~ date,
                y = ~ total,
                type = 'scatter', 
                mode = 'none', 
                groupnorm = 'percent',
                color = ~ continent_name,
                stackgroup = 'one') %>%
  plotly::layout(yaxis = list(title = "",
                              showgrid = FALSE,
                              hoverformat = ".2f",
                              ticksuffix = "%"),
                 xaxis = list(title = "",
                              showgrid = FALSE), 
                 margin = list(r = 10, l = 10, t = 0, b = 10),
                 hovermode = "x unified")
```


About
===================================== 

### About

**The Coronavirus Dashboard** 

This is a supporting dashboard for the [coronavirus](https://github.com/RamiKrispin/coronavirus) R package. It provides an overview of the 2019 Novel Coronavirus COVID-19 (2019-nCoV) epidemic. This dashboard is built with R using the [Rmakrdown](https://rmarkdown.rstudio.com/) using [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) framework and can easily reproduce by others. The code behind the dashboard is available on this [repo](https://github.com/RamiKrispin/coronavirus_dashboard).


**Data**

The input data for this dashboard is the [coronavirus](https://github.com/RamiKrispin/coronavirus) R package (dev version). The data and dashboard is refreshed on a daily bases. The raw data pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [repository](https://github.com/RamiKrispin/coronavirus-csv)

**Packages**

* Dashboard interface - the [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) package. 
* Visualization - the [plotly](https://plot.ly/r/) package
* Data manipulation - [dplyr](https://dplyr.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/) packages

**License**

All the code under this dashboard is completely open-sourced under [MIT](https://opensource.org/licenses/MIT) license

**Deployment and reproducibly**

The dashboard was deployed to Github docs. If you wish to deploy and/or modify the dashboard on your Github account, you can apply the following steps:

* Fork the dashboard [repository](https://github.com/RamiKrispin/coronavirus_dashboard), or
* Clone it and push it to your Github package
* Here some general guidance about deployment of flexdashboard on Github page - [link](https://github.com/pbatey/flexdashboard-example)

For any question or feedback, you can either open an [issue](https://github.com/RamiKrispin/coronavirus_dashboard/issues) or contact me on [Twitter](https://twitter.com/Rami_Krispin).
